instal---
title: "WiSTEM UNESCO"
author: "James Welch, After iGEM Women in STEM"
date: "5/25/2021"
output:
  html_document: default
---
# META
[R guide](https://apiportal.uis.unesco.org/BDDS_R_Tutorial)

# Package and Data Import
```{r, results='hide'}
library(tidyverse)
```



```{r}
#setwd("C:/Users/James/Dropbox/iGEM/WiSTEM/UNESCO")


inno_country <- read_csv('INNO_COUNTRY.csv',na='', col_types = 'cc')
inno_data_nat <- read_csv('INNO_DATA_NATIONAL.csv',na='',col_types = 'ccddcc')
inno_label <- read_csv('INNO_LABEL.csv',na='',col_types = 'cc')
inno_meta <- read_csv('INNO_METADATA.csv',na='',col_types = 'ccdcc')

sci_country <- read_csv('SCI_COUNTRY.csv',na='',col_types = 'cc')
sci_data_nat <- read_csv('SCI_DATA_NATIONAL.csv',na='',col_types = 'ccnncc')
sci_data_reg <- read_csv('SCI_DATA_REGIONAL.csv',na='',col_types = 'ccnncc')
sci_label <- read_csv('SCI_LABEL.csv',na='',col_types = 'cc')
sci_meta <- read_csv('SCI_METADATA.csv',na='',col_types = 'ccncc')
sci_region <- read_csv('SCI_REGION.csv',na='',col_types = 'ccc')



```

# Understanding the information

#### sci/inno_label
- INDICATOR_ID: Indicator code
- INDICATOR_LABEL_EN: Indicator code English label

#### sci/inno_DATA_NATIONAL
- INDICATOR_ID: Indicator code
- COUNTRY_ID: ISO 3166-1 alpha-3 country code
- YEAR: Year of data collection
- VALUE: measured value
- MAGNITUDE: metadata describing value (See below)
- QUALIFIER: describes quality of data collection (see below)

#### sci_DATA_REGIONAL
- INDICATOR_ID: Indicator code
- REGION_ID: General Region
- YEAR: Year of data collection
- VALUE: measured value
- MAGNITUDE: metadata describing value (See below)
- QUALIFIER: describes quality of data collection (see below)

#### sci/inno_METADATA
- INDICATOR_ID: indicator code
- COUNTRY_ID: ISO 3166-1 alpha-3 country code
- YEAR: year of data collection
- TYPE: type of metadata
- METADATA: metadata value

#### sci/inno_COUNTRY
- COUNTRY_ID: ISO 3166-1 alpha-3 country code
- COUNTRY_NAME_EN: UNSTATS M49 STANDARD Country name

#### sci_REGION
- REGION_ID: Name (acronym) of org responsible for regional composition + ":" + region name.
- COUNTRY_ID: ISO 3166-1 alpha-3 country code
- COUNTRY_NAME_EN: UNSTATS M49 STANDARD Country name

#### METADATA Notes:
- INDICATOR_ID : formal definition found at [here](http://uis.unesco.org/en/glossary)

#### MAGNITUDE: 
- **NIL**: value is 0, actually empty.
- **NA**: value is 0, not applicable 
- **SUPP**: value is empty, suppressed by country
- **LOWREL**: value is numeric, low reliability
- **INCLUDED**: value is empty, included in another data point
- **INCLUDES**: value is numeric, includes another data point

#### QUALIFIER:
- **NAT_EST**: estimate produced by country
- **UIS_EST**: estimate produced by UNESCO




# Helper Function to subset data by supplied values.
```{r}
subsetData <- function(dataset, years, countries, indicatorcodes, region = FALSE){
  a_subset <- dataset %>% filter(.,
    YEAR %in% years,
    if(!region){COUNTRY_ID %in% countries
      }else{REGION_ID %in% countries},
    INDICATOR_ID %in% indicatorcodes
  )
  return(a_subset)
}
```


# Main Parsing Function

Takes data set choice, scale, a set of years, a set of countries, and a set of indicator codes. 
```{r}
easy_subset <- function(choice, scale = 'national', years, countries, indicatorcodes) {
  choice = str_to_lower(choice)
  scale  = str_to_lower(scale)
  
  data <- paste0(choice,"_data_",substr(scale,0,3))
  
  if(data == 'inno_data_reg'){
    stop("INNO and REGEION is not a valid combination")
  }
  if(substr(scale,0,3)=='reg'){
    region = TRUE
  }else { region = FALSE}
  
  data <- subsetData(get(data),years,countries,indicatorcodes, region = region)

  data <- data %>% left_join(., 
                             get(paste0(choice,"_label")),
                             by = 'INDICATOR_ID')
  
  if(scale == 'regional'){
    data <- data %>% left_join(.,get(paste0(choice,'_region')),by='REGION_ID')
  } else {
    data <- data %>% left_join(.,get(paste0(choice,'_country')), by='COUNTRY_ID')
  }
  
  cleaned_meta <- get(paste0(choice,'_meta')) %>% 
                    mutate(., type_data = paste(TYPE, "_", METADATA)) %>% 
                    group_by(., INDICATOR_ID, COUNTRY_ID, YEAR) %>% 
                    summarise(type_data = list(type_data))
  
  data <- data %>% left_join(.,cleaned_meta,by=c('INDICATOR_ID', 'COUNTRY_ID', 'YEAR'))
  
  
  return(data)

  }

```


# Create new dataframe with filters
```{r}
# Use unique(data$column) to get all values.
year_range = c(2005:2011) 
country_range = c('USA','ARG')
ind_code = unique(sci_label$INDICATOR_ID)

filtered_data <- easy_subset(choice = 'inno',           # "inno" or "sci"
                             scale = 'national',        # "national" or "regional"
                             years = year_range,        # Modify Above
                             countries = country_range, # Modify Above
                             indicatorcodes = ind_code) # Modify Above

```



# Export Data
```{r}
write_csv(filtered_data, 'UNESCO_filtered.csv')
```

